Inc COVID-19 Case Forecast Eval: Results

### 2023-12-18 Code/Manuscript_figures.Rmd
Update Manuscript_figures.Rmd

@@

load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))

@@

Fig1.B <-dat_US_state %>%
  filter(sub_date <= cut_date) %>%
  filter(location=="US", 
         model=="COVIDhub-4_week_ensemble") %>% 
  filter(sub_date %in% cast_dates$sub_date) %>%
  mutate(value = case_when(quantile==0.5 ~ round(value),
                             quantile<0.5 ~ floor(value),
                             quantile>0.5 ~ ceiling(value),
                             type=='point' ~ round(value)), 
         model=as.factor(model), 
         model=droplevels(model)) %>% 
  pivot_wider(names_from = c(type, quantile), values_from=value) %>%
  rename(point = point_NA) %>% 
  ungroup() %>%
  ggplot(.) +
  geom_line(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey85") +
  geom_point(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey75") +
  geom_line(aes(target_end_date, point, group=interaction(sub_date, model), color=model), size=1.05) + 
  geom_point(aes(target_end_date, point, color=model), size=1.25) +
  geom_ribbon(aes(x=target_end_date, ymin=quantile_0.025, ymax=quantile_0.975,
                  fill=model, group=interaction(sub_date, model)), alpha=0.25) +
  geom_vline(xintercept = as.numeric(cut_date),  linetype=4, size=1.05) +
  scale_colour_manual(name="",
                      values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
  scale_fill_manual(name="",
                    values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
   scale_x_date(date_labels="%b-%y", breaks = obs_data$date, date_breaks ="3 months", date_minor_breaks ="1 month") +
  labs(title="B. Select National forecasts",
       y="Incident COVID-19 Cases") +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5),
        legend.title = element_blank(), legend.position = c(0.20, 0.82)) + scale_y_continuous(label=comma,
                                                                                               breaks=pretty_breaks(2))
plot_grid(Fig1.A, Fig1.B,
          nrow=2)

@@

Figure 2. 95% coverage and relWIS  and submissions per team

@@

WIS.heatmap <-mean_WIS %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 1.0, na.value = "grey75") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="Relative Weighted Interval Score",
       title="B. Relative Weighted Interval Score") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), axis.text.y=element_blank(),
        axis.text=axis.text_standard,
        legend.position="bottom")

@@

cov.heatmap <-cov.95 %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "white", mid = "#f7fcb9", high = "#31a354", midpoint = 0.75, na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="95% Coverage",
       title="C. 95% Coverage") +
  theme(axis.title.x = element_blank(), #axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(), axis.text.y=element_blank(),
        axis.text=axis.text_standard,
        legend.position="bottom")

@@

all_subs <-subs_nat %>%
  bind_rows(., subs_state) %>% 
  mutate(model_reorder=factor(model, levels=levels_model), 
         key=factor(key, levels=levels_key),
         value_display=format(round(per_sub, digits=2), nsmall = 2),
         value_display=ifelse(value_display=="  NA", NA, value_display)) %>%
  bind_rows(., missings) %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=per_sub)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient(low = "#fde0dd", high = "#c51b8a", na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="% Weeks \nSubmitted",
       title="A. Percent of Weeks Submitted") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        axis.text=axis.text_standard,
        legend.position="bottom")

@@

# Figures
load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))

@@

# Largest Counties
load(here("Data/cases_large counties_forecasts_for analysis_2022-08-18.Rdata"))
coverage_count <-cov.fig_dat(filter(dat_large_count, sub_date <= cut_date))
labels_count <-data_labels(coverage_count) 
Cov.Co <-cov.plot(coverage_count, labels_count) +
  labs(title="C. Coverage for Large County Forecasts*",
       x="Expected Coverage")

@@

# Figure
relWIS$model_reorder <-fct_rev(relWIS$model_reorder)
ggplot(relWIS, aes(model_reorder, location_name)) +
  geom_tile(aes(fill=log_WIS)) +
  geom_text(aes(label=scaled_rel_skill), size=3) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 0, na.value = "grey75", 
                       name = "Relative\nWeighted Interval Score", 
                       breaks = c(-1,0,1,2), 
                       labels =c("0.5", 1, 2, 3)) + 
  theme_bw() +
  labs(fill="Relative Weighted Interval Score") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        axis.text=axis.text_standard, plot.caption = plot.caption_standard)

@@

load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))
load(here("Data/cases_large counties_forecasts_for analysis_2022-08-18.Rdata"))

@@

### Figure 7 . Marginal mean WIS

@@

  marg_means[[i]] <-ggemmeans(model[[i]], terms=c("phase")) %>% #, "cases_log_std[-2:2]")) %>%
    rename(phase=x,
           cases=group) %>%

@@

# Figure
Fig7A <-marg_means_int %>%
  ggplot(.) +
  geom_point(aes(model, predicted_back_rescale)) +
  geom_linerange(aes(model, ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale)) +
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_lo_rescale), color="red", linetype="dashed") + 
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_up_rescale), color="red", linetype="dashed") +
  labs( y="Marginal mean\n Weighted Interval Score", 
        title="A.") +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard, axis.text.y=axis.text_standard,
        legend.position = "bottom", legend.title = element_blank(), legend.text=legend.text_standard,
        strip.text=strip.text_standard, strip.background=strip.background_standard,
        axis.title.x = element_blank(),
        #axis.text.x = element_text(angle = 45, hjust=1),
        axis.text.x = element_blank())

@@

Fig7B <-marg_means %>%
  mutate(scaled_pred=predicted_back/baseline$predicted_back,
         SE_scaled=(predicted_back_up-predicted_back_lo)/3.92,
         prob_base=round(pnorm(baseline$predicted_back,
                               predicted_back, SE_scaled), 5),
         prob_label=ifelse(prob_base >=0.9, "*", NA),
         phase=fct_rev(phase)) %>%
  ggplot(.) +
  geom_tile(aes(model, phase, fill=scaled_pred)) +
  geom_text(aes(model, phase, label=prob_label), size=4) +
  scale_fill_gradient2(low = "#7fcdbb", high = "#fecc5c", midpoint = 1.0,
                       name="Marginal mean Weighted Interval Score  \nscaled to the marginal mean  \nof COVIDhub-baseline model  \nacross all phases",
                       guide = guide_colourbar(barheight = 1.5)) +
  labs(y="", title="B.",
       caption="*80% CI is lower than baseline") + 
  theme_bw() +
  theme(legend.position = "bottom", legend.title.align=0.5, legend.box.just = "center", 
    plot.caption = plot.caption_standard,
    strip.text=strip.text_standard, strip.background=strip.background_standard,
    panel.border = element_rect(colour = "black", fill = NA),
    axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1), axis.text=axis.text_standard,
    axis.ticks.y=element_blank())

@@

```{r Fig7, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=8.5}

plot_grid(Fig7A, Fig7B, rel_heights  = c(0.5, 1), axis = "l",
          align="v", nrow = 2)

@@

>>>>

heatmap_format <-theme(axis.title.x = element_blank(), 
                       axis.title.y = element_blank(), 
                       axis.text=axis.text_standard,
                       legend.position="bottom", legend.title.align=0.5, legend.box.just = "center")

@@

load(here("Data/cases_nat_state_forecasts_for analysis.Rdata"))

@@

Fig1.B <-dat_US_state %>%
  filter(sub_date <= cut_date) %>%
  filter(location=="US", 
         model=="COVIDhub-4_week_ensemble") %>% 
  filter(sub_date %in% cast_dates$sub_date) %>%
  mutate(value = case_when(quantile==0.5 ~ round(value),
                             quantile<0.5 ~ floor(value),
                             quantile>0.5 ~ ceiling(value),
                             type=='point' ~ round(value)), 
         model=as.factor(model), 
         model=droplevels(model)) %>% 
  pivot_wider(names_from = c(type, quantile), values_from=value) %>%
  rename(point = point_NA) %>% 
  ungroup() %>%
  ggplot(.) +
  geom_line(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey85") +
  geom_point(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey75") +
  geom_line(aes(target_end_date, point, group=interaction(sub_date, model), color=model), size=1.05) + 
  geom_point(aes(target_end_date, point, color=model), size=1.25) +
  geom_ribbon(aes(x=target_end_date, ymin=quantile_0.025, ymax=quantile_0.975,
                  fill=model, group=interaction(sub_date, model)), alpha=0.25) +
  geom_vline(xintercept = as.numeric(cut_date),  linetype=4, size=1.05) +
  scale_colour_manual(name="",
                      values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
  scale_fill_manual(name="",
                    values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
   scale_x_date(date_labels="%b-%y", breaks = obs_data$date, date_breaks ="3 months", date_minor_breaks ="1 month") +
  labs(title="B. Select National forecasts",
       y="Incident COVID-19 Cases") +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5),
        legend.title = element_blank(), legend.position = c(0.20, 0.82)) + 
  scale_y_continuous(label=comma, breaks=pretty_breaks(2))

plot_grid(Fig1.A, Fig1.B,
          nrow=2)

@@

### Figure 2. 95% coverage and relWIS  and submissions per team

@@

WIS.heatmap <-mean_WIS %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 1.0, na.value = "grey75") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="Relative Weighted Interval Score",
       title="B. Relative Weighted Interval Score") +
  heatmap_format + theme(axis.text.y=element_blank())

@@

cov.heatmap <-cov.95 %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "white", mid = "#f7fcb9", high = "#31a354", midpoint = 0.75, na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="95% Coverage",
       title="C. 95% Coverage") +
  heatmap_format + theme(axis.text.y=element_blank())

@@

all_subs <-subs_nat %>%
  bind_rows(., subs_state) %>% 
  mutate(model_reorder=factor(model, levels=levels_model), 
         key=factor(key, levels=levels_key),
         value_display=format(round(per_sub, digits=2), nsmall = 2),
         value_display=ifelse(value_display=="  NA", NA, value_display)) %>%
  bind_rows(., missings) %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=per_sub)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient(low = "#fde0dd", high = "#c51b8a", na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="% Weeks \nSubmitted",
       title="A. Percent of Weeks Submitted") +
  heatmap_format

@@

# Figures
load(here("Data/cases_nat_state_forecasts_for analysis.Rdata"))

@@

# Largest Counties
load(here("Data/cases_large counties_forecasts_for analysis.Rdata"))
coverage_count <-cov.fig_dat(filter(dat_large_count, sub_date <= cut_date))
labels_count <-data_labels(coverage_count) 
Cov.Co <-cov.plot(coverage_count, labels_count) +
  labs(title="C. Coverage for Large County Forecasts*",
       x="Expected Coverage")

@@

# Figure
relWIS$model_reorder <-fct_rev(relWIS$model_reorder)
ggplot(relWIS, aes(model_reorder, location_name)) +
  geom_tile(aes(fill=log_WIS)) +
  geom_text(aes(label=scaled_rel_skill), size=3) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 0, na.value = "grey75", 
                       name = "Relative\nWeighted Interval Score", 
                       breaks = c(-1,0,1,2), 
                       labels =c("0.5", 1, 2, 3)) + 
  theme_bw() +
  labs(fill="Relative Weighted Interval Score") +
  heatmap_format +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        plot.caption = plot.caption_standard)

@@

load(here("Data/cases_nat_state_forecasts_for analysis.Rdata"))
load(here("Data/cases_large counties_forecasts_for analysis.Rdata"))

@@

### Figure 7 . Marginal mean WIS and 95% COverage

@@

  marg_means[[i]] <-as.data.frame(ggemmeans(model[[i]], terms=c("phase")) %>% #, "cases_log_std[-2:2]")) %>%
    rename(phase=x,
           cases=group) 
    )

  marg_means[[i]] <-marg_means[[i]] %>%

@@

# Figure
Fig7A <-marg_means_int %>%
  ggplot(.) +
  geom_point(aes(model, predicted_back_rescale)) +
  geom_linerange(aes(model, ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale)) +
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_lo_rescale), color="red", linetype="dashed") + 
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_up_rescale), color="red", linetype="dashed") +
  labs( y="Marginal mean\n Weighted Interval Score", 
        title="A.") +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard, axis.text.y=axis.text_standard,
        axis.title.x = element_blank(),
        axis.text.x = element_blank())

@@

Fig7B <-marg_means %>%
  mutate(scaled_pred=predicted_back/baseline$predicted_back,
         SE_scaled=(predicted_back_up-predicted_back_lo)/3.92,
         prob_base=round(pnorm(baseline$predicted_back,
                               predicted_back, SE_scaled), 5),
         prob_label=ifelse(prob_base >=0.9, "*", NA),
         phase=fct_rev(phase)) %>%
  ggplot(.) +
  geom_tile(aes(model, phase, fill=scaled_pred)) +
  geom_text(aes(model, phase, label=prob_label), size=4) +
  scale_fill_gradient2(low = "#7fcdbb", high = "#fecc5c", midpoint = 1.0,
                       name="Marginal mean Weighted Interval Score  \nscaled to the marginal mean  \nof COVIDhub-baseline model  \nacross all phases",
                       guide = guide_colourbar(barheight = 1.5)) +
  labs(y="", title="B.",
       caption="*80% CI is lower than baseline") + 
  theme_bw() +
  heatmap_format +
  theme(axis.text.x = element_blank(), #axis.text.x = element_text(angle = 45, hjust=1), 
      axis.ticks.y=element_blank())

@@

load(here("Data/WIS_all horizons_nat_state.rdata")) # data
load("./Data/covid_epi_phases.rdata")

Fig7C <-WIS_all %>%
  filter(location!="US", 
         range==95)%>%
  dplyr::select( "location_name", "location", "target", "date", "model" ,
                 "sub_date","coverage") %>% 
  distinct() %>%
  left_join(., phases, by=c("date", "location_name", "location")) %>%
  drop_na() %>%
  mutate(phase=str_to_title(phase)) %>%
  mutate(model=as.factor(model),
         model=factor(model, levels = levels),
         phase=factor(phase, 
                      levels = c("Increasing", "Peak", "Decreasing",  "Nadir")),
         phase=fct_rev(phase)
  ) %>%
  group_by(phase, model) %>%
  summarise(cov=round(mean(coverage), 2)) %>%
  ggplot() +
  geom_tile(aes(model, phase, fill=cov)) +
  geom_text(aes(model, phase, label=cov), size=2.5) +
  scale_fill_gradient2(low = "white", mid = "#f7fcb9", high = "#31a354", midpoint = 0.75, na.value = "grey75",
                       name="95% Prediction Interval Coverage",
                       guide = guide_colourbar(barheight = 1.5)) +
  labs(y="", title="C.") + 
  theme_bw() +
  heatmap_format +
  theme(panel.border = element_rect(colour = "black", fill = NA),
        axis.text.x = element_text(angle = 45, hjust=1), 
        axis.ticks.y=element_blank())

@@

```{r Fig7, echo=FALSE, warning=F, message=F, fig.height=11, fig.width=8.5}

plot_grid(Fig7A, Fig7B, Fig7C,
          rel_heights  = c(0.35, 0.50, 0.70), axis = "l",
          align="v", nrow = 3)

### Figure 8: % Prediction in Each Direction

#levels <-levels[!levels %in% "COVIDhub-baseline"]
#
dat <-dat_US_state %>%
  filter(#!grepl("baseline", model), 
         quantile==0.25 | quantile==0.75) %>%
  pivot_wider(names_from = c(type, quantile), values_from=value)

base <-dat_US_state %>%
  filter(grepl("baseline", model),
         quantile==0.5) %>%
  pivot_wider(names_from = c(type, quantile), values_from=value) %>%
  rename(quantile_0.50_base=quantile_0.5) %>%
  dplyr::select(-model, -sub_date, -forecast_date)

direct <-dat %>%
  left_join(., base, by=c("location", "target", "target_end_date")) %>%
  mutate(pred_class=as.factor(ifelse(quantile_0.25 > quantile_0.50_base, "increasing",
                      ifelse(quantile_0.75 < quantile_0.50_base, "decreasing", "stable/uncertain")))) %>%
  rename(date=target_end_date) %>%
  left_join(., phases, by=c("date", "location")) %>%
  drop_na() %>%
  mutate(model=as.factor(model), 
         model=factor(model, levels=levels),
         phase=as.factor(str_to_title(phase)),
         phase=factor(phase, levels=c("Nadir", "Decreasing", "Peak", "Increasing")))

team_phase_pred <-crossing(direct$model, direct$pred_class, direct$phase) %>%
  rename(model=1, 
         pred_class=2,
         phase=3) #%>%
  #filter(!grepl("baseline", model))

team_phase <-direct %>%
  group_by(model, phase) %>%
  count() %>%
  rename(phase_denom=n)

direct <-direct %>%
  group_by(model, phase, pred_class) %>%
  count(.) %>%
  ungroup() %>%
  right_join(., team_phase_pred, by=c("model", "phase", "pred_class")) %>%
  left_join(., team_phase, by=c("model", "phase")) %>%
  mutate(precent_pred=n/phase_denom, 
         precent_pred=ifelse(is.na(precent_pred), 0, precent_pred), 
         value_display=format(round(precent_pred, digits=2), nsmall = 2),
         value_display=ifelse(value_display=="  NA", NA, value_display))

Fig8A <-direct %>%
  filter(pred_class=="increasing") %>%
  ggplot(aes(model, phase)) +
  geom_tile(aes(fill=precent_pred)) +
  geom_text(aes(label=value_display), size=2.75) +
  scale_fill_gradient2(low = "#fdd0a2", high = "#e6550d",  na.value = "grey75",
                       name = stringr::str_wrap("Percent of forecasts with predicted increasing trajectory",
                                                 width = 35), 
                       guide = guide_colourbar(barheight = 1.5, title.position = "top")) +#, title.hjust=0.5)) +
  labs(y="", title="A.") + 
  theme_bw() +
  heatmap_format +
  theme(axis.text.x = element_blank(), 
        axis.ticks.y=element_blank(), 
        legend.position = "right", legend.direction = "horizontal", legend.title.align=0)

Fig8B <-direct %>%
  filter(pred_class=="stable/uncertain") %>%
  ggplot(aes(model, phase)) +
  geom_tile(aes(fill=precent_pred)) +
  geom_text(aes(label=value_display), size=2.75) +
  scale_fill_gradient2(low = "#dadaeb", high = "#54278f",  na.value = "grey75",
                       name = stringr::str_wrap("Percent of forecasts with predicted stable or uncertain trajectory",
                                                 width = 35), 
                       guide = guide_colourbar(barheight = 1.5, title.position = "top")) +#, title.hjust=0.5)) +
  labs(y="", title="B.") + 
  theme_bw() +
  heatmap_format +
  theme(axis.text.x = element_blank(), 
        axis.ticks.y=element_blank(), 
        legend.position = "right", legend.direction = "horizontal", legend.title.align=0)

Fig8C <-direct %>%
  filter(pred_class=="decreasing") %>%
  ggplot(aes(model, phase)) +
  geom_tile(aes(fill=precent_pred)) +
  geom_text(aes(label=value_display), size=2.75) +
  scale_fill_gradient2(low = "#a1d99b", high = "#006d2c", na.value = "grey75",
                       name = stringr::str_wrap("Percent of forecasts with predicted decreasing trajectory",
                                                 width = 35), 
                       guide = guide_colourbar(barheight = 1.5, title.position = "top")) +#, title.hjust=0.5)) +
  labs(y="", title="C.") + 
  theme_bw() +
  heatmap_format +
  theme(axis.text.x = element_text(angle = 45, hjust=1), 
        axis.ticks.y=element_blank(), 
        legend.position = "right", legend.direction = "horizontal", legend.title.align=0)

```{r Fig9, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=11}

plot_grid(Fig8A, Fig8B, Fig8C,
          rel_heights  = c(0.50, 0.50, 0.82), axis = "l",
          align="v", nrow = 3)
